The landscape of chloroplast genome assembly tools

### 2019-05-17 bmc_article.tex
Update on Overleaf.

@@

%%% Put your definitions there:
\startlocaldefs
\newcommand{\formatprogramnames}[1]{\texttt{#1}}
\newcommand{\ce}{\formatprogramnames{chloroExtractor}}
\newcommand{\oa}{\formatprogramnames{ORG.Asm}}
\newcommand{\fp}{\formatprogramnames{Fast-Plast}}
\newcommand{\ioga}{\formatprogramnames{IOGA}}
\newcommand{\np}{\formatprogramnames{NOVOPlasty}}
\newcommand{\go}{\formatprogramnames{GetOrganelle}}
\newcommand{\cassp}{\formatprogramnames{Chloroplast assembly protocol}}
\newcommand{\docker}[2]{\href{https://cloud.docker.com/u/chloroextractorteam/repository/docker/chloroextractorteam/#1}{chloroextracteam/#1:#2}}
\newcommand{\dockerce}{\docker{benchmark\_chloroextractor}{v2.0.0}}
\newcommand{\dockercesha}{sha256:8f49e03424b37e699c5fea9391f79dbb9f2d0dc550ce86c4c700f39deec2dacd}
\endlocaldefs

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Genome sequencing - important bilogical insights.
Eukaryotes have organelles with separate genomes.
Particularly, plants have chloroplasts beside mitochondria (generally chloroplast or plastid? plastid bit more general, I think also chromoplasts etc).
Chloroplast genome content and structure reviewed elsewhere \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses.
Comparative genomics even more informative.
Complete chloroplast genome much easier to get than complete core genome.
But specific genome structure (esp. inverted repeat) makes automated resolution with short reads hard.
Many genome sequencing project contain chloroplast reads as a by-product even though that was not the aim of the study.
Resolved chloroplast helps to remove these reads which can improve assembly of the core.
Databases contain short read data for species where no reference chloroplast is available in the databases.
There is a lot of potential use cases for chloroplast genomes \cite{tonti-filippini_what_2017}
Resolving these would allow for more and larger scale comparative studies on whole chloroplasts.
Open question (?) structural variation across taxonomic groups.
Some researchers use shallow genome sequencing to reconstruct full chloroplast genomes for use as super-barcodes \cite{coissac_barcodes_2016}.

\subsection*{Chloroplast genome architecture}
Maybe separate section describing a prototypical chloroplast: LSC, IRa, SSC, IRb. Gene number and content. Size ranges. Evolutionary history incl. endosymbiontic gene transfer. Reduced genomes in parasitic plants. Loss of ndh gene family (maybe too specific). Preliminary SNP analysis in A. thaliana to estimate intraspecific variability? Is there anything known about intra-organismic variability (do some plants have multiple populations of chloroplasts? probably off topic).

@@

\section*{Methods}
Source code for all methods at \url{https://github.com/chloroExtractorTeam/benchmark} (zenodo doi), docker containers at ..., datasets at (zenodo links for simulated and real).
This study adheres to the guidelines for computational method benchmarking \cite{weber_essential_2018}.

@@

\subsection*{Data}
\subsubsection*{Simulated}
A. thaliana (TAIR10). Perfect data, sliding window with seqkit (circular for chloro and mito). No errors, no insert size variation, perfect quality, equal coverage.

@@

These questions were answered with "GOOD", "OK" or "BAD", depending on the quality of the result. 
For example, a "GOOD" installation utilizes a automated package or dependency management like apt,
CRAN, docker, etc. An "OK" installation procedure provides a custom script to install everything
or at least lists all dependencies. A "BAD" installation procedure fails to list important dependencies
or produces errors, that prevent a successful installation.

@@

\section*{Results}
\subsection*{Computational Resources}
Table and figure of performance evaluation. Including RAM, I/O, runtime, parallelization ability, CPU usage. Free text description of results, range, mean/median, effect of input size, threads parameter, ...
\subsection*{Quantitative}

\subsection*{Qualitative}

\begin{table}[]
    \centering
    \caption{Overview of the results of the qualitative usability evaluation. Each tool 
    could score "GOOD", "OK" or "BAD" in each of the categories.}
    \label{tab:resultsQual}
\begin{tabular}{lcccccc}   
Tool & Installation & Test/Tutorial & Documentation & Maintenance & FLOSS\\                           \hline 
\go{}    & OK   & OK   & GOOD & GOOD & GOOD\\  
\fp{}      & BAD  & OK   & GOOD & GOOD & GOOD\\ 
\oa{}         & BAD  & BAD  & OK   & GOOD & GOOD\\ 
\np{}      & GOOD & GOOD & GOOD & GOOD & OK\\
\ce{} & GOOD & GOOD & GOOD & GOOD & GOOD\\
\ioga{}            & BAD  & BAD  & OK   & OK   & BAD\\
\cassp{}    & OK   & GOOD & OK   & GOOD & GOOD\\ 
\end{tabular}      
\end{table}

Most of the tools were evaluated as mainly "GOOD" (see table \ref{tab:resultsQual}). However, a few critique points remained.
Two minor dependencies were missing in the \go{} installation instructions. There was also no test data available and an issue occurred when running it on a \textit{A.~thaliana} dataset. We are currently in the process of resolving this with the authors.

The \fp{ installation instructions were missing some dependencies. Like \go{}, \fp{} does not offer a test dataset or a tutorial, except from some example commands.

@@

\subsection*{Quantitative}
\subsubsection*{Simulated data}
\textit{tile plot for simulated data goes here}
The only assembler obtaining perfect results according to our score for the simulated datasets is \go{}. \ioga{} and CAP showed the worst performance being unable to fully assemble a single chloroplast out of 14. \np{} performed second best with scores above 80 for all datasets, only failing to resolve the chloroplast contig into one single circular assembly. As could have been expected the overall performance is best, when the input data data consists purely of chloroplast reads. Only \ioga{} and CAP failed to deliver any results once each. Besides that no clear causality between length of the input reads, ratio of core vs chloroplast reads and the performance can be clearly observed.

\subsubsection*{real data sets}
\textit{swarm plot goes here}
%% Den upset plot eventuell erst in die Discussion reinbringen ? Da wir davon ableiten welche tools wir empfehlen?

@@

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}

>>>>

\usepackage{cleveref}

\usepackage{biocon}
\newplant{At}{genus=Arabidopsis,epithet=thaliana}

@@

%%% Put your definitions there:
\startlocaldefs
\newcommand{\formatprogramnames}[1]{\texttt{#1}}
\newcommand{\ce}{\formatprogramnames{chloroExtractor}}
\newcommand{\oa}{\formatprogramnames{ORG.Asm}}
\newcommand{\fp}{\formatprogramnames{Fast-Plast}}
\newcommand{\ioga}{\formatprogramnames{IOGA}}
\newcommand{\np}{\formatprogramnames{NOVOPlasty}}
\newcommand{\go}{\formatprogramnames{GetOrganelle}}
\newcommand{\cassp}{\formatprogramnames{Chloroplast assembly protocol}}

% Niklas table
%\newcommand{\ok}{\textcolor[rgb]{0.7,0.7,0}{\textsc{\bfseries{}OKAY}}}
%\newcommand{\bad}{\textcolor{red}{\textsc{\bfseries{}BAD}}}
%\newcommand{\good}{\textcolor[rgb]{0,0.6,0}{\textsc{\bfseries{}GOOD}}}
\newcommand{\ok}{\textsc{Okay}}
\newcommand{\bad}{\textsc{Bad}}
\newcommand{\good}{\textsc{Good}}

% List of docker images
\newcommand{\docker}[2]{\href{https://cloud.docker.com/u/chloroextractorteam/repository/docker/chloroextractorteam/#1}{chloroextracteam/#1:#2}}
\newcommand{\dockerce}{\docker{benchmark\_chloroextractor}{v2.0.0}}
\newcommand{\dockercesha}{\texttt{8f49e03424b37e699c5fea9391f79dbb9f2d0dc550ce86c4c700f39deec2dacd}}
\newcommand{\dockeroa}{\docker{benchmark\_org-asm}{v2.0.0}}
\newcommand{\dockeroasha}{\texttt{ff83677c97b7c4e346191b9e04a5162eeea2a8dfd9e60ff84067d33747868f60}}
\newcommand{\dockerfp}{\docker{benchmark\_fastplast}{v2.0.0}}
\newcommand{\dockerfpsha}{\texttt{a3d06a610f8340ba49c3ff3b27342534e2a5348b17caa4fd3c3f3d327243a272}}
\newcommand{\dockerioga}{\docker{benchmark\_ioga}{v2.0.0}}
\newcommand{\dockeriogasha}{\texttt{4698a21c343c60290bbf16811165a654ff01e8fddeb41cd75d0771b7f45968c0}}
\newcommand{\dockernp}{\docker{benchmark\_novoplasty}{v2.0.0}}
\newcommand{\dockernpsha}{\texttt{106387bad4e8e5c53fb9d4c3bdd60cdddf05a47b52a38697369eeabb947c547c}}
\newcommand{\dockergo}{\docker{benchmark\_getorganelle}{v2.0.0}}
\newcommand{\dockergosha}{\texttt{2ed3a464a82025a196ea56c649d0b0a3472cef76994bb065d56054d93298d956}}
\newcommand{\dockercassp}{\docker{benchmark\_chloroplast\_assembly\_protocol}{v2.0.0}}
\newcommand{\dockercasspsha}{\texttt{d10d2317103d91fadec4cb1d5466ac109f23176ad74b6f44f0179175f050155d}}
\endlocaldefs

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Genome sequencing - important biological insights.
Eukaryotes have organelles with separate genomes.
Particularly, plants have chloroplasts beside mitochondria (generally chloroplast or plastid? plastid bit more general, I think also chromoplasts etc).
Chloroplast genome content and structure reviewed elsewhere \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses.
Comparative genomics even more informative.
Complete chloroplast genome much easier to get than complete core genome.
But specific genome structure (esp. inverted repeat) makes automated resolution with short reads hard.
Many genome sequencing project contain chloroplast reads as a by-product even though that was not the aim of the study.
Resolved chloroplast helps to remove these reads which can improve assembly of the core.
Databases contain short read data for species where no reference chloroplast is available in the databases.
There is a lot of potential use cases for chloroplast genomes \cite{tonti-filippini_what_2017}
Resolving these would allow for more and larger scale comparative studies on whole chloroplasts.
Open question (?) structural variation across taxonomic groups.
Some researchers use shallow genome sequencing to reconstruct full chloroplast genomes for use as super-barcodes \cite{coissac_barcodes_2016}.

\subsection*{Chloroplast genome architecture}
Maybe separate section describing a prototypical chloroplast: LSC, IRa, SSC, IRb. Gene number and content. Size ranges. Evolutionary history incl. endosymbiontic gene transfer. Reduced genomes in parasitic plants. Loss of ndh gene family (maybe too specific). Preliminary SNP analysis in \plant{At} to estimate intraspecific variability? Is there anything known about intra-organismic variability (do some plants have multiple populations of chloroplasts? probably off topic).

@@

\section*{Methods}
Source code for all methods is available at \url{https://github.com/chloroExtractorTeam/benchmark} and archived under (zenodo doi).
All docker images are published on \url{https://cloud.docker.com/u/chloroextractorteam/} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).
datasets at (zenodo links for simulated and real).
This study adheres to the guidelines for computational method benchmarking \cite{weber_essential_2018}.

@@

\subsection*{Data}
\subsubsection*{Simulated}
\plant{At} (TAIR10). Perfect data, sliding window with seqkit (circular for chloro and mito). No errors, no insert size variation, perfect quality, equal coverage.

@@

These questions were answered with \good{}, \ok{} or \bad{}, depending on the quality of the result. 
For example, a \good{} installation utilizes a automated package or dependency management like apt,
CRAN, docker, etc. An \ok{} installation procedure provides a custom script to install everything
or at least lists all dependencies. A \bad{} installation procedure fails to list important dependencies
or produces errors, that prevent a successful installation.

@@

\section*{Results}
\subsection*{Computational Resources}
Table and figure of performance evaluation. Including RAM, I/O, runtime, parallelization ability, CPU usage. Free text description of results, range, mean/median, effect of input size, threads parameter, ...

\subsection*{Qualitative}
Most of the tools were evaluated as mainly \good{} (see table \ref{tab:resultsQual}). However, a few critique points remained.
Two minor dependencies were missing in the \go{} installation instructions. There was also no test data available and an issue occurred when running it on a \plant{At} dataset. We are currently in the process of resolving this with the authors.

The \fp{} installation instructions were missing some dependencies. Like \go{}, \fp{} does not offer a test dataset or a tutorial, except from some example commands.

@@

\subsection*{Quantitative}
\subsubsection*{Simulated data}
\textit{tile plot for simulated data goes here}
The only assembler obtaining perfect results according to our score for the simulated datasets is \go{}.
\ioga{} and \cassp{} showed the worst performance being unable to fully assemble a single chloroplast out of 14.
\np{} performed second best with scores above 80 for all datasets, only failing to resolve the chloroplast contig into one single circular assembly.
(As could have been expected -> discussion) the overall performance is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results once each.
Besides that no clear causality between length of the input reads, ratio of core vs chloroplast reads and the performance can be clearly observed.

\subsubsection*{real data sets}
Highest scoring is \go{} next is \fp{} while \ioga{} and \cassp{} perform poorly, never reaching a score of 1 (figure 1). TODO: median/mean score, mad/sd Tabelle

\textit{swarm plot goes here}
%% Den upset plot eventuell erst in die Discussion reinbringen ? Da wir davon ableiten welche tools wir empfehlen?

@@

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{plots/real_datasets_tree.pdf}
  \caption{\csentence{Sample figure title.}
      Figure legend text.}
      \end{figure}

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[]
    \centering
    \caption{Overview of the results of the qualitative usability evaluation. Each tool 
    could score \good{}, \ok{} or \bad{} in each of the categories.}
    \label{tab:resultsQual}
\begin{tabular}{p{3cm}cccccc}   
Tool & Installation & Test/Tutorial & Documentation & Maintenance & FLOSS\\                           \hline \ce{}     &  \good{}  &  \good{}  &  \good{}  &  \good{}  &  \good{} \\
\cassp{}  &  \ok{}    &  \good{}  &  \ok{}    &  \good{}  &  \good{} \\
\fp{}     &  \bad{}   &  \ok{}    &  \good{}  &  \good{}  &  \good{} \\
\go{}     &  \ok{}    &  \ok{}    &  \good{}  &  \good{}  &  \good{} \\
\ioga{}   &  \bad{}   &  \bad{}   &  \ok{}    &  \ok{}    &  \bad{}  \\
\np{}     &  \good{}  &  \good{}  &  \good{}  &  \good{}  &  \ok{}   \\
\oa{}     &  \bad{}   &  \bad{}   &  \ok{}    &  \good{}  &  \good{} \\ \hline
\end{tabular}      
\end{table}

\begin{table}[h!]
\caption{Docker images used in our benchmark setup.}
\label{tab:dockerimages}
\centering
      \resizebox{\textwidth}{!}{\begin{tabular}{ccc}
        \hline
          Tool & Image name and tag & SHA256 Checksum   \\ \hline
          \ce & \dockerce & \dockercesha \\
          \cassp & \dockercassp & \dockercasspsha \\
          \fp & \dockerfp & \dockerfpsha \\
          \go & \dockergo & \dockergosha \\
          \ioga & \dockerioga & \dockeriogasha \\
          \np & \dockernp & \dockernpsha \\
          \oa & \dockeroa & \dockeroasha \\ \hline
      \end{tabular}}
\end{table}

\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}
